Continuous Deductive Reasoning over Streamed RDF Data

### 2015-12-02 article.tex
fixed incorrect interval notations and inconsistent variable names

However, this definition of a C-Datalog program does not take into
account any notion of windowing over the stream $S$, and so any fact of
which an instance occurs in $S$ will be considered true from the instant
of its occurrence in $S$.  While this is technically acceptable in the
case of sets of rules without constraints, if a constraint exists and its
body becomes satisfied by some $\Pi^C$ at some instant $i$ then the
program will be unsatisfiable from that instant on, making $E^S$ empty
from then on.  As such, we define a {\em Windowed Continuous Datalog program}
$\Pi^{C, r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in (1,s) \} \rangle\} \rangle$
(where $r = max(r_1,\dots,r_s)$) to be a tuple with components $R$ and $F^0$ as defined above for a
C-Datalog program, but with each part of $S = \mathop{\cup}^s_{x=1}S_x$
associated with the range $r_x \in \mathbb{N}$ of a sliding window over
$S_x$, for $x \in (1,s)$. This specifies that the fact instances
from $S_x$ are to contribute to the entailments of $\Pi^{C,r}$ from the instant
they occur in $S_x$ for a duration of $r_x$ consecutive instants.

@@

We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0,
\{ W^{S_x, r_x} \mid x \in (1,s) \} \rangle$
at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
applying the rules $R$ to the union of the persistent facts $F^0$ and all
finite states $W^{S_x,r_x}_i$ of windows $W^{S_x,r_x}$ over streams $S_x$ of range
$r_x \in \mathbb{N}$, for all $x \in (1,s)$; we refer to this as the {\em instantaneous state} of
the WC-Datalog program, written
$\Pi^{C,r}_i = \langle R, F^0 \cup \mathop{\cup}\limits_{x=1}^s W^{S_x,r_x}_i \rangle$.
As previously shown in \refdef{continuous datalog: ta datalog program}, this may then be further
reduced to the Datalog program
$\Pi = \langle R, strip(F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_x,r_x}_i \rangle$.

@@

  With regards to the notation of WC-Datalog, these built-in predicates
  may be represented by the set of facts $B$, or the set of persistent
  fact instances ${B^0 = \{ \langle b, 0 \rangle \mid b \in B \}}$, that
  are the collected extension of those predicates. A RIF-Core program may
  therefore be represented as a Datalog program
  ${\Pi = \langle R, F \cup B \rangle}$, a Time-Annotated RIF-Core program
  as a TA-Datalog program ${\Pi^{T} = \langle R, F^T \cup B^0 \rangle}$, a
  Continuous RIF-Core program as a C-Datalog program
  ${\Pi^{C} = \langle R, F^0 \cup B^0, S_1 \cup \dots \cup S_n \rangle}$
  and a Windowed Continuous RIF-Core program as a WC-Datalog program
  ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,n)\} \rangle}$.
  Furthermore, ${\langle R, B \rangle \models \varnothing}$ and
  ${\langle R, B^0 \rangle \models \varnothing}$, for any set $R$ of
  \emph{safe} RIF-Core rules. As such, the form of entailment for each
  type of RIF-Core program is the same as that for the equivalent variety
  of Datalog program:
  \begin{align*}
    \Pi =&& \langle R, F \cup B \rangle &\models E \\
    \Pi^{T} =&& \langle R, F^T \cup B^0 \rangle &\models E^T \\
    \Pi^{C} =&& \langle R, F^0 \cup B^0, S \rangle &\models E^S \\
    \Pi^{C,r} =&& \langle R, F^0 \cup B^0, S, r \rangle &\models \langle E^0, P, N \rangle \\
    \Pi^{C,r}_i =&& \langle R, F^0 \cup B^0 \cup W^{S,r}_i \rangle &\models E^0 \cup E^T_i
  \end{align*}

@@

\begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
  \emph{CoDeR} is an abstract model for \emph{Co}ntinuous \emph{De}ductive \emph{R}easoning over streamed RDF data that expresses the semantics of Windowed Continuous RIF-Core, thus supporting continuous reasoning with general knowledge expressed in both RIF-Core and OWL 2 RL.
  It is composed of a stream-based semantic data model and a minimal algebra of stream-to-stream operators, the former for expressing the persistent fact instances $F^0$ and $E^0$, the streams $S_x$, $P$ and $N$ and the window ranges $r_x$ for $x \in (1,s)$ of each $\Pi^{C,r}$, the latter for expressing the persistent set of rules $R$, the built-in predicates of RIF-Core $B^0$ and enforcing the window ranges $r_x$.
  It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.

  % Stream-based Data Model
  \begin{nestedsection}{Data Model}{model: data}
    The data model of CoDeR is composed of multiple-source multiple-sink data streams, as shown in \reffig{CoDeR: stream}.
  The data model of CoDeR is composed of multiple-source multiple-sink data streams, as shown in \reffig{CoDeR: stream}.
  These streams are based on the notion of in-order streams presented in \refdef{continuous datalog: stream}, but are divided into two classes: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
  In addition, the notion of an \emph{instance} is extended in \refdef{CoDeR: graph instance} from that of an instance of a \emph{fact} that became true at a specific instant, as presented in \refdef{continuous datalog: fact instance}.
  The streams of CoDeR are streams of this extended form of instance, as per \refdef{CoDeR: stream}, which are compatible with the ongoing work on continuous query semantics by the RSP working group\footnote{According to \url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}, 4th Nov 2015, a CoDeR graph instance may be expressed as an RSP graph with two timestamps predicated as `became valid' and `ceased to be valid'.}.
  \begin{definition}[Graph Instance]\labeldef{CoDeR: graph instance}
    A \emph{graph instance} is an RDF graph with some additional temporal annotation regarding both the instant from which it became valid and the instant at which it will cease to be valid;
    it is a triple composed of the graph ${\{(s,p,o),\dots,(s',p',o')\}}$ and the pair of instants $e$ and $n$ that bound the interval ${(e,n]}$ for which the graph is valid, written ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
    The names $e$ and $n$ are so named because they relate to the instants of the instance's {$e$ntailment} and {$n$egation} by the program.
    It should be noted that validity is a stronger concept than truth, as all facts within a graph that is valid in a system are understood to be true, but facts may be true despite not being represented in a graph that is valid within the system.
    Two graph instances containing the same facts and annotation for $e$ and $n$ are considered identical;
    two graph instances containing the same facts but different annotations for $e$ and/or $n$ are considered to be distinct.
  \end{definition}
  \begin{definition}[CoDeR Stream]\labeldef{CoDeR: stream}
    All \emph{streams} of CoDeR are sequences of graph instances, using the definition of \emph{sequence} used in \refdef{continuous datalog: stream} of streams in WC-Datalog.
    They are also \emph{in-order} with regards to their temporal annotation $e$, but potentially \emph{out-of-order} with regards to their temporal annotation $n$.
    In practise, where a stream received from some source external to the model is not annotated with one or either, these instants are to be assigned by some trivial pre-processing;
    this will be based on either the existing annotation, if it exists, or the current time, and the sliding window definitions in the modelled $\Pi^{C,r}$.
  \end{definition}
  \begin{definition}[Empty Stream]\labeldef{CoDeR: empty stream}
    The \emph{empty stream} is a CoDeR stream that contains no graph instances, and is denoted by the symbol $\varnothing$.
    This means that the stream will contain no graph instances that map to instant $i$, for all instants ${i \in \mathbb{T}}$.
    This stream may occur in specific expressions of any program when the range of an operator is mutually exclusive to that of the operator that succeeds it in a dataflow.
  \end{definition}

@@

  As has been noted in \refdef{CoDeR: graph instance}, the interval ${(e,n]}$ for any graph instance is the the period of time for which that graph instance is \emph{valid} \citep{SemanticStreamingManagement,sparkwave}.
  The notion of validity within WC-RIF-Core programs relates to the period of time for which a given fact instance ${\langle f, t \rangle}$ is justified by a state ${\Pi^{C,r}_i = \langle R, F^0 \cup B^0 \cup \mathop{\cup}^s_{x=1} W^{S_x,r_x}_i \rangle \models \langle f, t \rangle}$ of a program $\Pi^{C,r}$.
  As such, the interval ${(e,n]}$ is used to express the window ranges $r_x$ for $x \in (1,s)$ of the ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,s) \} \rangle}$:
  for graph instances entering a CoDeR-based system, $e$ is the instant at which every fact in the given graph was asserted, and $n$ is the instant at which they will leave the semantic window over $S_x$ described by $\Pi^{C,r}$ (a.k.a. being ``negated''), where ${n = e + r_x}$ in the general case;
  for graph instances produced by a CoDeR-based system, $e$ is the latest instant at which any fact within the given graph was asserted, and $n$ is the earliest instant at which any fact within the graph will be negated.
  \reffig{CoDeR: intersected-intervals,CoDeR: consecutive states} provide an example of this for the ${\Pi^{C,3} = \langle \{c \Leftarrow a \land b\},\{\},\{ W^{S,3} \}, 3\rangle}$, where ${S = \langle a,1 \rangle,\langle b,2 \rangle}$ over instants ${i = 1 \dots 4}$.
  \begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{intersected-intervals-example}
    \caption[Derivation of valid-time for entailed fact instance from those of its justification]{Derivation of valid-time for entailed fact instance \emph{c} from those of its justification \emph{a} and \emph{b}.}
    \labelfig{CoDeR: intersected-intervals}
  \end{figure}
  \begin{figure}[h!]
    \centering
    \begin{align*}
      \Pi^{C,3}_{1} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle \} \rangle & \models & \{ \} \\
      \Pi^{C,3}_{2} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
      \Pi^{C,3}_{3} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
      \Pi^{C,3}_{4} & = \langle \{c \Leftarrow a \land b\}, \{ \langle b, 2 \rangle \} & \models & \{ \}
    \end{align*}
    \caption{Consecutive states of a WC-RIF-Core program and their entailments}
    \labelfig{CoDeR: consecutive states}
  \end{figure}

  The first class of CoDeR streams, input/output streams, provide the semantics of the streams and sets $S_x$, $P$, $N$, $F^0$ and $E^0$, for any ${\Pi^{C,r} = \langle R,F^0,\{W^{S_x,r_x} \mid x \in (1,s)\} \rangle \models \langle E^0, P, N \rangle}$ expressed by a CoDeR-based system.
  These streams and sets are represented as CoDeR streams, as per \refdef{CoDeR: stream}, that are, in the case of input streams, produced elsewhere in the Semantic Web or, in the case of output streams, exposed on the Semantic Web as a resource themselves, as per \refdef{CoDeR: input/output}.
  \begin{definition}[Input/Output Stream]\labeldef{CoDeR: input/output}
    Both \emph{input} and \emph{output} streams of CoDeR are CoDeR streams that are identified on the Semantic Web by IRIs.
    The graphs of which there are instances in an input/output stream are heterogeneous, in so much as each graph may contain any number of triples, each in turn being composed of any values.
  \end{definition}
  In each CoDeR-based system the input streams express the logical streams $S_x$ for $x \in (1,s)$ of the given $\Pi^{C,r}$, and a single output stream expresses the logical stream $P$ entailed by the $\Pi^{C,r}$.
  This single output stream also expresses the stream $N$ entailed by the $\Pi^{C,r}$, as that logical stream is defined as the set of instances in the stream $P$ ordered in a sequence by the instant $n$ at which ${\langle R, F^0 \cup B^0 \cup \mathop{\cup}^s_{x=1} W^{S_x,r_x}_{n}}$ \emph{ceased} to justify them;
  this instant $n$ is, therefore, the same $n$ with which each graph is annotated in CoDeR, being the instant at which the graph instance will cease to be valid.
  As such, the annotation of streamed RDF graphs with the interval ${(e,n]}$ gives \emph{all} streams of CoDeR a dual semantics with regard to WC-RIF-Core.

@@

    The algebra fundamentally consists of the following set of stream-to-stream operators: ${\{{}_g\pi_t, \sigma_{triple}, \sigma_{external}, \coloneq_{head}, \cup, \rstreamjoin\}}$.
    These operators encapsulate the practical expression of the set of functions required to derive the entailment of WC-RIF-Core programs in the context of CoDeR input/output streams.
    These functions are those required for WC-Datalog programs, with the addition of two functions:
    a one-pass unary evaluation function $\sigma_{external}$ for the built-in predicates of RIF-Core;
    and a one-pass unary graph decomposition function ${{}_g\pi_t}$ to convert the CoDeR input stream of heterogeneous RDF graph instances into a stream of homogeneous RDF triple instances, the latter being the form of a stream $S_x$ of a program ${\Pi^{C,r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in (1,n) \} \rangle}$.
    The continuous derivation of the entailment of WC-Datalog programs requires, in turn, the set of functions required for querying deductive databases using Datalog, with the addition of another two functions:
    a two-pass unary valid-window maintenance function $w$ to enforce the windows of range $r_x$, for $x \in (1,n)$, over the CoDeR input streams;
    and a two-pass unary window-to-stream function $s$ to extract the stream of novel graph instances from a sequence of consecutive valid-windows, closely related to the ${IStream()}$ function of CQL \citep{StanfordStream}.
    Finally, the functions required for querying deductive databases using Datalog are conventionally the complete set required of the relational algebra\footnote{This refers to the complete relational algebra less the functionality of set minus $\setminus$, which is analogous to the NaF operator $\sim$ that is not supported by positive Datalog.}, being ${\{\sigma,\pi,\Join,\cup\}}$ as described by \citet{codd70relationalModel}.
    However, three of these may be restricted when processing semantic data, becoming:
    the one-pass unary equality-based triple-select $\sigma_{triple}$;
    the two-pass binary natural join $\Join^\ast$ between \emph{sets} of variable bindings for graph patterns (requiring valid-windows over streams as input, and producing valid-windows);
    and the one-pass unary template-project $\pi_{triple}$ that projects constant values and bindings for variables in graph patterns from graph instances into new triple instances.
    The one-pass binary set operator $\cup$ utilised by the relational algebra need only be modified to preserve the mapping of instances to instants when applied to CoDeR streams.

@@

  \begin{nestedsection}{Summary of CoDeR}{CoDeR: summary}
    Given the time-based nature of windowing and of the justification of entailments in WC-RIF-Core, it is possible to determine in advance the instant at which fact instances will be negated.
    Furthermore, facts in RDF may be grouped into graphs, meaning that fact instances with the same annotation that will be negated at the same instant may be grouped into an annotated graph without loss of information.
    As such, both streams $P$ and $N$ entailed by a $\Pi^{C,r}$ may be represented by streams of CoDeR's twice-annotated graphs.
    As the SteM $\rstreamjoin$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
    Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
    as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the states of a given WC-RIF-Core program ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,n) \} \rangle}$.

>>>>

However, this definition of a C-Datalog program does not take into
account any notion of windowing over the stream $S$, and so any fact of
which an instance occurs in $S$ will be considered true from the instant
of its occurrence in $S$.  While this is technically acceptable in the
case of sets of rules without constraints, if a constraint exists and its
body becomes satisfied by some $\Pi^C$ at some instant $i$ then the
program will be unsatisfiable from that instant on, making $E^S$ empty
from then on.  As such, we define a {\em Windowed Continuous Datalog program}
$\Pi^{C, r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in [1,s] \} \rangle\} \rangle$
(where $r = max(r_1,\dots,r_s)$) to be a tuple with components $R$ and $F^0$ as defined above for a
C-Datalog program, but with each part of $S = \mathop{\cup}^s_{x=1}S_x$
associated with the range $r_x \in \mathbb{N}$ of a sliding window over
$S_x$, for $x \in [1,s]$. This specifies that the fact instances
from $S_x$ are to contribute to the entailments of $\Pi^{C,r}$ from the instant
they occur in $S_x$ for a duration of $r_x$ consecutive instants.

@@

We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0,
\{ W^{S_x, r_x} \mid x \in [1,s] \} \rangle$
at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
applying the rules $R$ to the union of the persistent facts $F^0$ and all
finite states $W^{S_x,r_x}_i$ of windows $W^{S_x,r_x}$ over streams $S_x$ of range
$r_x \in \mathbb{N}$, for all $x \in [1,s]$; we refer to this as the {\em instantaneous state} of
the WC-Datalog program, written
$\Pi^{C,r}_i = \langle R, F^0 \cup \mathop{\cup}\limits_{x=1}^s W^{S_x,r_x}_i \rangle$.
As previously shown in \refdef{continuous datalog: ta datalog program}, this may then be further
reduced to the Datalog program
$\Pi = \langle R, strip(F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_x,r_x}_i \rangle$.

@@

  With regards to the notation of WC-Datalog, these built-in predicates
  may be represented by the set of facts $B$, or the set of persistent
  fact instances ${B^0 = \{ \langle b, 0 \rangle \mid b \in B \}}$, that
  are the collected extension of those predicates. A RIF-Core program may
  therefore be represented as a Datalog program
  ${\Pi = \langle R, F \cup B \rangle}$, a Time-Annotated RIF-Core program
  as a TA-Datalog program ${\Pi^{T} = \langle R, F^T \cup B^0 \rangle}$, a
  Continuous RIF-Core program as a C-Datalog program
  ${\Pi^{C} = \langle R, F^0 \cup B^0, S_1 \cup \dots \cup S_n \rangle}$
  and a Windowed Continuous RIF-Core program as a WC-Datalog program
  ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in [1,s]\} \rangle}$.
  Furthermore, ${\langle R, B \rangle \models \varnothing}$ and
  ${\langle R, B^0 \rangle \models \varnothing}$, for any set $R$ of
  \emph{safe} RIF-Core rules. As such, the form of entailment for each
  type of RIF-Core program is the same as that for the equivalent variety
  of Datalog program:
  \begin{align*}
    \Pi =&& \langle R, F \cup B \rangle &\models E \\
    \Pi^{T} =&& \langle R, F^T \cup B^0 \rangle &\models E^T \\
    \Pi^{C} =&& \langle R, F^0 \cup B^0, S \rangle &\models E^S \\
    \Pi^{C,r} =&& \langle R, F^0 \cup B^0, S, r \rangle &\models \langle E^0, P, N \rangle \\
    \Pi^{C,r}_i =&& \langle R, F^0 \cup B^0 \cup W^{S,r}_i \rangle &\models E^0 \cup E^T_i
  \end{align*}

@@

\begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
  \emph{CoDeR} is an abstract model for \emph{Co}ntinuous \emph{De}ductive \emph{R}easoning over streamed RDF data that expresses the semantics of Windowed Continuous RIF-Core, thus supporting continuous reasoning with general knowledge expressed in both RIF-Core and OWL 2 RL.
  It is composed of a stream-based semantic data model and a minimal algebra of stream-to-stream operators, the former for expressing the persistent fact instances $F^0$ and $E^0$, the streams $S_x$, $P$ and $N$ and the window ranges $r_x$ for $x \in [1,s]$ of each $\Pi^{C,r}$, the latter for expressing the persistent set of rules $R$, the built-in predicates of RIF-Core $B^0$ and enforcing the window ranges $r_x$.
  It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.

  % Stream-based Data Model
  \begin{nestedsection}{Data Model}{model: data}
    The data model of CoDeR is composed of multiple-source multiple-sink data streams, as shown in \reffig{CoDeR: stream}.
  The data model of CoDeR is composed of multiple-source multiple-sink data streams, as shown in \reffig{CoDeR: stream}.
  These streams are based on the notion of in-order streams presented in \refdef{continuous datalog: stream}, but are divided into two classes: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
  In addition, the notion of an \emph{instance} is extended in \refdef{CoDeR: graph instance} from that of an instance of a \emph{fact} that became true at a specific instant, as presented in \refdef{continuous datalog: fact instance}.
  The streams of CoDeR are streams of this extended form of instance, as per \refdef{CoDeR: stream}, which are compatible with the ongoing work on continuous query semantics by the RSP working group\footnote{According to \url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}, 4th Nov 2015, a CoDeR graph instance may be expressed as an RSP graph with two timestamps predicated as `became valid' and `ceased to be valid'.}.
  \begin{definition}[Graph Instance]\labeldef{CoDeR: graph instance}
    A \emph{graph instance} is an RDF graph with some additional temporal annotation regarding both the instant from which it became valid and the instant at which it will cease to be valid;
    it is a triple composed of the graph ${\{(s,p,o),\dots,(s',p',o')\}}$ and the pair of instants $e$ and $n$ that bound the interval ${[e,n)}$ for which the graph is valid, written ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
    The names $e$ and $n$ are so named because they relate to the instants of the instance's {$e$ntailment} and {$n$egation} by the program.
    It should be noted that validity is a stronger concept than truth, as all facts within a graph that is valid in a system are understood to be true, but facts may be true despite not being represented in a graph that is valid within the system.
    Two graph instances containing the same facts and annotation for $e$ and $n$ are considered identical;
    two graph instances containing the same facts but different annotations for $e$ and/or $n$ are considered to be distinct.
  \end{definition}
  \begin{definition}[CoDeR Stream]\labeldef{CoDeR: stream}
    All \emph{streams} of CoDeR are sequences of graph instances, using the definition of \emph{sequence} used in \refdef{continuous datalog: stream} of streams in WC-Datalog.
    They are also \emph{in-order} with regards to their temporal annotation $e$, but potentially \emph{out-of-order} with regards to their temporal annotation $n$.
    In practise, where a stream received from some source external to the model is not annotated with one or either, these instants are to be assigned by some trivial pre-processing;
    this will be based on either the existing annotation, if it exists, or the current time, and the sliding window definitions in the modelled $\Pi^{C,r}$.
  \end{definition}
  \begin{definition}[Empty Stream]\labeldef{CoDeR: empty stream}
    The \emph{empty stream} is a CoDeR stream that contains no graph instances, and is denoted by the symbol $\varnothing$.
    This means that the stream will contain no graph instances that map to instant $i$, for all instants ${i \in \mathbb{T}}$.
    This stream may occur in specific expressions of any program when the range of an operator is mutually exclusive to that of the operator that succeeds it in a dataflow.
  \end{definition}

@@

  As has been noted in \refdef{CoDeR: graph instance}, the interval ${[e,n)}$ for any graph instance is the the period of time for which that graph instance is \emph{valid} \citep{SemanticStreamingManagement,sparkwave}.
  The notion of validity within WC-RIF-Core programs relates to the period of time for which a given fact instance ${\langle f, t \rangle}$ is justified by a state ${\Pi^{C,r}_i = \langle R, F^0 \cup B^0 \cup \mathop{\cup}^s_{x=1} W^{S_x,r_x}_i \rangle \models \langle f, t \rangle}$ of a program $\Pi^{C,r}$.
  As such, the interval ${[e,n)}$ is used to express the window ranges $r_x$ for $x \in [1,s]$ of the ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in [1,s] \} \rangle}$:
  for graph instances entering a CoDeR-based system, $e$ is the instant at which every fact in the given graph was asserted, and $n$ is the instant at which they will leave the semantic window over $S_x$ described by $\Pi^{C,r}$ (a.k.a. being ``negated''), where ${n = e + r_x}$ in the general case;
  for graph instances produced by a CoDeR-based system, $e$ is the latest instant at which any fact within the given graph was asserted, and $n$ is the earliest instant at which any fact within the graph will be negated.
  \reffig{CoDeR: intersected-intervals,CoDeR: consecutive states} provide an example of this for the ${\Pi^{C,3} = \langle \{c \Leftarrow a \land b\},\{\},\{ W^{S,3} \}, 3\rangle}$, where ${S = \langle a,1 \rangle,\langle b,2 \rangle}$ over instants ${i = 1 \dots 4}$.
  \begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{intersected-intervals-example}
    \caption[Derivation of valid-time for entailed fact instance from those of its justification]{Derivation of valid-time for entailed fact instance \emph{c} from those of its justification \emph{a} and \emph{b}.}
    \labelfig{CoDeR: intersected-intervals}
  \end{figure}
  \begin{figure}[h!]
    \centering
    \begin{align*}
      \Pi^{C,3}_{1} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle \} \rangle & \models & \{ \} \\
      \Pi^{C,3}_{2} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
      \Pi^{C,3}_{3} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
      \Pi^{C,3}_{4} & = \langle \{c \Leftarrow a \land b\}, \{ \langle b, 2 \rangle \} & \models & \{ \}
    \end{align*}
    \caption{Consecutive states of a WC-RIF-Core program and their entailments}
    \labelfig{CoDeR: consecutive states}
  \end{figure}

  The first class of CoDeR streams, input/output streams, provide the semantics of the streams and sets $S_x$, $P$, $N$, $F^0$ and $E^0$, for any ${\Pi^{C,r} = \langle R,F^0,\{W^{S_x,r_x} \mid x \in [1,s]\} \rangle \models \langle E^0, P, N \rangle}$ expressed by a CoDeR-based system.
  These streams and sets are represented as CoDeR streams, as per \refdef{CoDeR: stream}, that are, in the case of input streams, produced elsewhere in the Semantic Web or, in the case of output streams, exposed on the Semantic Web as a resource themselves, as per \refdef{CoDeR: input/output}.
  \begin{definition}[Input/Output Stream]\labeldef{CoDeR: input/output}
    Both \emph{input} and \emph{output} streams of CoDeR are CoDeR streams that are identified on the Semantic Web by IRIs.
    The graphs of which there are instances in an input/output stream are heterogeneous, in so much as each graph may contain any number of triples, each in turn being composed of any values.
  \end{definition}
  In each CoDeR-based system the input streams express the logical streams $S_x$ for $x \in [1,s]$ of the given $\Pi^{C,r}$, and a single output stream expresses the logical stream $P$ entailed by the $\Pi^{C,r}$.
  This single output stream also expresses the stream $N$ entailed by the $\Pi^{C,r}$, as that logical stream is defined as the set of instances in the stream $P$ ordered in a sequence by the instant $n$ at which ${\langle R, F^0 \cup B^0 \cup \mathop{\cup}^s_{x=1} W^{S_x,r_x}_{n}}$ \emph{ceased} to justify them;
  this instant $n$ is, therefore, the same $n$ with which each graph is annotated in CoDeR, being the instant at which the graph instance will cease to be valid.
  As such, the annotation of streamed RDF graphs with the interval ${[e,n)}$ gives \emph{all} streams of CoDeR a dual semantics with regard to WC-RIF-Core.

@@

    The algebra fundamentally consists of the following set of stream-to-stream operators: ${\{{}_g\pi_t, \sigma_{triple}, \sigma_{external}, \coloneq_{head}, \cup, \rstreamjoin\}}$.
    These operators encapsulate the practical expression of the set of functions required to derive the entailment of WC-RIF-Core programs in the context of CoDeR input/output streams.
    These functions are those required for WC-Datalog programs, with the addition of two functions:
    a one-pass unary evaluation function $\sigma_{external}$ for the built-in predicates of RIF-Core;
    and a one-pass unary graph decomposition function ${{}_g\pi_t}$ to convert the CoDeR input stream of heterogeneous RDF graph instances into a stream of homogeneous RDF triple instances, the latter being the form of a stream $S_x$ of a program ${\Pi^{C,r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in [1,s] \} \rangle}$.
    The continuous derivation of the entailment of WC-Datalog programs requires, in turn, the set of functions required for querying deductive databases using Datalog, with the addition of another two functions:
    a two-pass unary valid-window maintenance function $w$ to enforce the windows of range $r_x$, for $x \in [1,s]$, over the CoDeR input streams;
    and a two-pass unary window-to-stream function $s$ to extract the stream of novel graph instances from a sequence of consecutive valid-windows, closely related to the ${IStream()}$ function of CQL \citep{StanfordStream}.
    Finally, the functions required for querying deductive databases using Datalog are conventionally the complete set required of the relational algebra\footnote{This refers to the complete relational algebra less the functionality of set minus $\setminus$, which is analogous to the NaF operator $\sim$ that is not supported by positive Datalog.}, being ${\{\sigma,\pi,\Join,\cup\}}$ as described by \citet{codd70relationalModel}.
    However, three of these may be restricted when processing semantic data, becoming:
    the one-pass unary equality-based triple-select $\sigma_{triple}$;
    the two-pass binary natural join $\Join^\ast$ between \emph{sets} of variable bindings for graph patterns (requiring valid-windows over streams as input, and producing valid-windows);
    and the one-pass unary template-project $\pi_{triple}$ that projects constant values and bindings for variables in graph patterns from graph instances into new triple instances.
    The one-pass binary set operator $\cup$ utilised by the relational algebra need only be modified to preserve the mapping of instances to instants when applied to CoDeR streams.

@@

  \begin{nestedsection}{Summary of CoDeR}{CoDeR: summary}
    Given the time-based nature of windowing and of the justification of entailments in WC-RIF-Core, it is possible to determine in advance the instant at which fact instances will be negated.
    Furthermore, facts in RDF may be grouped into graphs, meaning that fact instances with the same annotation that will be negated at the same instant may be grouped into an annotated graph without loss of information.
    As such, both streams $P$ and $N$ entailed by a $\Pi^{C,r}$ may be represented by streams of CoDeR's twice-annotated graphs.
    As the SteM $\rstreamjoin$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
    Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
    as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the states of a given WC-RIF-Core program ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in [1,s] \} \rangle}$.
